% !TEX ROOT = ../Thesis.tex 
%-
%-> 中文摘要
%-
\chapter{摘\quad 要}\chaptermark{摘\quad 要}% 摘要标题
% \setcounter{page}{1}% 开始页码
% \pagenumbering{Roman}% 页码符号
% 22p / 12p = 1.83
\linespread{1.5}
\zihao{-4}
% \setlength{\baselineskip}{20pt}

在“双碳”战略持续推进和交通运输体系加速电气化的背景下，电动汽车进入规模化应用阶段。动力电池与旋转机械作为支撑整车动力输出的关键部件，在长期复杂工况下易发生性能退化与结构故障，直接影响车辆运行安全、系统可靠性与服役寿命。依托多源传感数据构建数据驱动的智能预测与健康管理方法，已成为保障电动汽车全生命周期高可靠运行的重要技术路径。然而，受观测噪声、工况变化、域分布漂移和未知输入等因素影响，现有深度模型在实际部署中普遍存在推断偏差、分布外泛化能力不足以及大语言模型幻觉等问题，难以满足安全关键场景下的高可信决策需求。针对上述问题，本文以无偏不确定性量化为主线，面向电动汽车关键部件开展退化建模与故障诊断研究，构建覆盖可信状态估计、分布外鲁棒退化建模与抗幻觉故障诊断的智能预测与健康管理方法体系，
% 
本文的具体贡献如下：

（1）针对锂离子电池健康状态可信估计问题，提出了一种基于层次化随机时空 Transformer 的健康状态估计方法。
首次，该方法利用 Gumbel-Softmax 重参数化构建随机自注意力机制，使特征聚合过程具备不确定性量化能力。
其次，结合计及不确定性的逐时间步卷积、距离敏感的层次化 Transformer 编码器，实现时空不确定性的量化和时空退化特征的提取。
最后，利用基于保序回归的不确定性校准策略，实现时空退化特征的有效提取与预测置信度的无偏修正。
实验结果表明，所提出方法在多个锂离子电池循环老化数据集上均取得优于对比方法的估计精度与不确定性量化效果，并能够有效辨识分布内与分布外退化模式。

（2）针对锂离子电池在未知工况与未见退化模式下的泛化退化建模问题，提出了一种基于拓扑约束下不确定性驱动特征调弈的分布外鲁棒退化建模方法。
首先，该方法通过启发式循环信号重构与基于 Mamba 的流形收缩特征提取策略，构建低本征维度且具有拓扑连续性的退化表征。
其次，结合多元高斯统计建模与不确定性驱动特征合成策略，对潜在分布外退化模式进行刻画，
最后，设计长尾感知的自适应重平衡优化机制，以缓解样本分布不均引发的建模偏置。
基于多个开源电池数据集构建的分布外泛化实验表明，所提出方法在退化建模任务上表现出更强的鲁棒性与泛化能力。

（3）针对多模态大语言模型在旋转机械故障诊断中的幻觉问题，提出了一种基于扩散增强可信多模态大语言模型的抗幻觉故障诊断方法。
首先，该方法构建 ControlNet 引导的跨数据集条件扩散模型，生成多样化少样本故障信号，以缓解类别统计偏差。
其次，设计文本与信号对齐的可信多模态大语言模型，并通过关键 token 定位与不确定性重加权策略实现序列级幻觉检测。
最后，构建面向非预期输入的抗幻觉诊断流程，提升模型对未知输入的拒识能力与可靠诊断能力。
实验结果表明，所提出方法在多个故障诊断数据集和多项评价指标上均优于现有对比方法。

本文围绕电动汽车关键部件在复杂工况下的高可靠运行需求，系统研究无偏不确定性量化在可信估计、分布外泛化与抗幻觉诊断中的作用机理与实现方法。相关研究不仅提升数据驱动模型的预测精度与可靠性，也为构建面向实际应用场景的电动汽车智能预测与健康管理系统提供理论依据与方法支撑。

\keywords{电动汽车; 预测与健康管理; 深度学习; 锂离子电池; 旋转机械}% 中文关键词

%-
%-> 英文摘要
%-
\chapter{Abstract}\chaptermark{Abstract}% 摘要标题

With the continued advancement of the carbon peaking and carbon neutrality strategy and the accelerated electrification of the transportation system, electric vehicles have entered the stage of large-scale deployment. As key components that sustain the vehicle powertrain output, power batteries and rotating machinery are prone to performance degradation and structural faults under long-term and complex operating conditions, thereby directly affecting operational safety, system reliability, and service life. Constructing data-driven intelligent prognostics and health management methods based on multi-source sensing data has become an important technical pathway for ensuring the highly reliable operation of electric vehicles throughout their full life cycle. However, due to the effects of observation noise, operating condition variations, domain distribution shifts, and unknown inputs, existing deep models commonly suffer from inference bias, insufficient out-of-distribution generalization capability, and hallucinations in large language models during practical deployment, making them inadequate for high-confidence decision-making in safety-critical scenarios. To address these issues, this dissertation takes unbiased uncertainty quantification as the main theme and investigates degradation modeling and fault diagnosis for key electric vehicle components, thereby establishing an intelligent prognostics and health management framework covering trustworthy state estimation, out-of-distribution robust degradation modeling, and hallucination-resistant fault diagnosis.

The main contributions of this dissertation are as follows:

(1) To address the problem of trustworthy state-of-health estimation for lithium-ion batteries, a state-of-health estimation method based on a hierarchical stochastic spatiotemporal Transformer is proposed. The proposed method constructs a stochastic self-attention mechanism through Gumbel-Softmax reparameterization, enabling uncertainty quantification during feature aggregation. It further integrates uncertainty-aware step-wise convolution, a distance-aware hierarchical Transformer encoder, and an uncertainty calibration strategy based on isotonic regression to achieve effective extraction of spatiotemporal degradation features and unbiased correction of predictive confidence. Experimental results show that the proposed method achieves superior estimation accuracy and uncertainty quantification performance over competing methods on multiple lithium-ion battery cycling aging datasets, while effectively distinguishing in-distribution and out-of-distribution degradation patterns.

(2) To address the generalized degradation modeling problem of lithium-ion batteries under unknown operating conditions and unseen degradation patterns, an out-of-distribution robust degradation modeling method based on topology-constrained uncertainty-driven feature adaptation is proposed. The proposed method constructs degradation representations with low intrinsic dimensionality and topological continuity through heuristic cycling signal reconstruction and a Mamba-based manifold-shrinking feature extraction strategy. It further characterizes potential out-of-distribution degradation patterns by combining multivariate Gaussian statistical modeling with an uncertainty-driven feature synthesis strategy, and designs a long-tail-aware adaptive rebalancing optimization mechanism to alleviate modeling bias caused by imbalanced sample distributions. Out-of-distribution generalization experiments conducted on multiple open-source battery datasets demonstrate that the proposed method exhibits stronger robustness and generalization capability in degradation modeling tasks.

(3) To address hallucinations in multimodal large language models for rotating machinery fault diagnosis, a hallucination-resistant fault diagnosis method based on a diffusion-enhanced trustworthy multimodal large language model is proposed. The proposed method first constructs a ControlNet-guided cross-dataset conditional diffusion model to generate diverse few-shot fault signals, thereby alleviating category-level statistical bias. It then develops a trustworthy multimodal large language model with text-signal alignment and realizes sequence-level hallucination detection through key token localization and an uncertainty reweighting strategy. In addition, an anti-hallucination diagnostic pipeline for unexpected inputs is developed to improve the model's rejection capability and diagnostic reliability in the presence of unknown inputs. Experimental results demonstrate that the proposed method outperforms existing comparative methods on multiple fault diagnosis datasets and evaluation metrics.

In summary, this dissertation systematically investigates the underlying mechanisms and implementation methods of unbiased uncertainty quantification in trustworthy estimation, out-of-distribution generalization, and hallucination-resistant diagnosis for the highly reliable operation of key electric vehicle components under complex operating conditions. The reported studies not only improve the predictive accuracy and reliability of data-driven models, but also provide theoretical foundations and methodological support for building intelligent prognostics and health management systems for electric vehicles in practical application scenarios.

\englishkeywords{Electric vehicles; Prognostics and health management; Deep learning; Lithium-ion batteries; Rotating machinery; Large language models}
