\chapter[致谢]{致\quad 谢}\chaptermark{致\quad 谢}\label{chap:acknowledgements}
% syntax: \chapter[目录]{标题}\chaptermark{页眉}
% \thispagestyle{noheaderstyle}% 如果需要移除当前页的页眉
% \pagestyle{noheaderstyle}% 如果需要移除整章的页眉

\reviewORprint{
    % 盲审版
    感谢各位审稿人百忙之中拨冗审阅本文，并提出宝贵意见和中肯建议，您们投入的宝贵时间与专业反馈极大促进了本研究工作的完善与提升。
}{
    % 打印版
    攻读硕士期间

    感谢好师兄付饶博士
    
    感谢父亲、母亲和家人的
    
    感谢各位审稿人百忙之中拨冗审阅本文，并提出宝贵意见和中肯建议，您们投入的宝贵时间与专业反馈极大促进了本研究工作的完善与提升。
}

\chapter{攻读硕士学位期间取得的学术成果和获得的奖励}\label{chap:academic_achievements_and_awards}

\reviewORprint{
    % 盲审版
    
}{
    % 打印版

    % \section*{学术论文:}
    % \begin{enumerate}[leftmargin=*]
    %     \item Authors. Title[J]Journal, Year, Volume(Number):Pages.
    % \end{enumerate}
    
    \section*{第一作者发表/录用学术论文:}
    \begin{enumerate}[leftmargin=*]
        \item Hierarchical Stochastic Spatial-Temporal Transformer for Trustworthy State-of-Health Estimation of Batteries in Industrial Applications[J]. IEEE Transactions on Industrial Informatics, 2025, 21(11):9069-9080. (\textbf{SCI, JCR 1 区, 论文第二章})
    \end{enumerate}
    \section*{第一作者在审学术论文:}
    \begin{enumerate}[leftmargin=*]
        \item OOD-Robust Degradation Modeling for Vehicular Battery Management Systems via Uncertainty-Driven Feature Manipulation with Topological Constraint[J]. IEEE Transactions on Vehicular Technology, Submitted for Publication. (\textbf{SCI, JCR 1 区, 论文第三章})
        \item Hallucination-Resilient Fault Diagnosis of Rolling Machinery via Diffusion-Augmented Trustworthy Multimodal Large Language Model[J]. IEEE Transactions on Neural Networks and Learning Systems, Submitted for Publication. (\textbf{SCI, JCR 1 区, 论文第四章})
    \end{enumerate}
    \section*{合作作者发表/录用学术论文:}
    \begin{enumerate}[leftmargin=*]
        \item FedProLN: Class Prototype-Enhanced Federated Learning for Long-Tailed Noisy Labels[C]//Proc. IEEE ICASSP, Barcelona, Spain, 2026: xxx-xxx. (\textbf{CCF B, 第三作者})
        % \item Title[J]Journal, Year, Volume(Number):Pages. (\textbf{SCI, JCR1, 第四作者})
        % \item Title[J]Journal, Year, Volume(Number):Pages. (\textbf{SCI, JCR1, 第四作者})
    \end{enumerate}
    \section*{合作作者在审学术论文:}
    \begin{enumerate}[leftmargin=*]
        \item Learning Probabilistic Cross-Spatiotemporal Dependencies for Non-Stationary SOH Prediction of Lithium-Ion Batteries[J]. IEEE Transactions on Industrial Informatics, Submitted for Publication. (\textbf{SCI, JCR 1 区, 第四作者})
        \item Mixture-of-Experts Guided Siamese Mamba Model for Long-Term Degradation Prediction of Li-ion Batteries under Threefold Coupled Shifts[J]. IEEE Transactions on Intelligent Transportation Systems, Submitted for Publication. (\textbf{SCI, JCR 1 区, 第四作者})
        \item LEAF-Net: Balanced Time–Frequency Learning for City-Level Cellular Traffic Prediction[J]. IEEE Transactions on Mobile Computing, Submitted for Publication. (\textbf{SCI, JCR 1 区, 第三作者})
    \end{enumerate}
    \section*{专利:}
    \begin{enumerate}[leftmargin=*]
        \item 毕远国, 林新辉, 付饶. 一种具有无偏不确定性量化能力的电池健康状态估计方法. 中国发明专利，
        \item 毕远国, 林新辉, 张志豪, 付饶. 一种面向分布漂移的电池跨域健康状态估计方法. 中国发明专利，
        \item 毕远国, 张志豪, 林新辉. 一种面向异构传感器数据缺失的锂离子电池荷电状态估计方法. 中国发明专利，
    \end{enumerate}
    \section*{获得的奖励:}
    \begin{enumerate}[leftmargin=*]
        \item 2025年度~研究生国家奖学金.
        \item 2025年度~东北大学优秀毕业生.
        \item 2025年度~东北大学优秀研究生.
        \item 2025年度~东北大学硕士研究生一等学业奖学金.
        \item 2024年度~东北大学硕士研究生一等学业奖学金.
        \item 2023年度~东北大学硕士研究生一等学业奖学金.
    \end{enumerate}
}

\cleardoublepage[plain]% 让文档总是结束于偶数页，可根据需要设定页眉页脚样式，如 [noheaderstyle]
